### 2020-11-09 paper/manuscript.tex
add tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required

%
\RequirePackage{fix-cm}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{moreverb,url}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[misc]{ifsym}
\usepackage{bm}
\usepackage{amsmath}  
\usepackage{float}
\usepackage{multirow}
\usepackage{threeparttable}  
\usepackage{booktabs}
\usepackage{color}
\definecolor{revision_switch}{RGB}{255,0,0}
\graphicspath{{./fig/}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Journal of Geodesy}
%
\begin{document}
\linenumbers
\title{GNSS displacement detection and extraction based on Bayesian inference
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}

%\titlerunning{Short form of title}        % if too long for running head

\author{Nan Shen\textsuperscript{1}\and
		Liang Chen\textsuperscript{12*}\and        
		Ruizhi Chen\textsuperscript{12} %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{
	\Letter Liang Chen\\
	{l.chen@whu.edu.cn}\\       
	\at
    {1} State Key Laboratory of Information Engineering in Surveying, Mapping and Remote Sensing, Wuhan University, 129 Luoyu Road, Wuhan 430079, China
    \at
    {2} Collaborative Innovation Center for Geospatial Technology,  129 Luoyu Road, Wuhan 430079, China
    \\
}

\date{Received: date }
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}

\keywords{GNSS\and unmodeled error\and machine learning\and convolutional neural network\and wavelet coherence}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}

\section{Methodology}
\subsection{GNSS Kinematic Positioning}

\subsection{Displacement Detection}
\begin{equation}\label{eq_ts}
\left[ {{x_1},{x_2}, \cdots {x_m}} \right]
\end{equation}
\begin{equation}\label{eq_cps}
\left[ {{\tau _1},{\tau _2}, \cdots {\tau _n}} \right]
\end{equation}
\begin{equation}\label{eq_ts_cps}
x \sim \left\{ {\begin{array}{*{20}{r}}
	{N({\mu _0},\sigma ),}&{t < {\tau _1}}\\
	{N({\mu _1},\sigma ),}&{{\tau _1} \le t < {\tau _2}}\\
	\vdots &{}\\
	{\begin{array}{*{20}{c}}
		{N({\mu _n},\sigma ),}
		\end{array}}&{{\tau _{n - 1}} \le t < {\tau _n}}\\
	{N({\mu _{n + 1}},\sigma ),}&{{\tau _n} \le t}
	\end{array}} \right.
\end{equation}

\section{Experiments and Results}

\section{Discussion}

\section{Conclusions}

\begin{acknowledgements}
This research was funded by the National Key Research and Development Programs 2018YFB0505400, the Natural Science Fund of Hubei Province 2018CFA007.
\end{acknowledgements}

%
\section*{Conflict of Interest}
The authors declare that they have no conflict of interest.
\section*{Author Contributions}
Conceptualization: Nan Shen; Methodology: Nan Shen; Writing - original draft preparation: Nan Shen; Writing - review and editing:Ruizhi Chen; Supervision: Liang Chen.
 \section*{Data Availability}

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{ref}

\end{document}